Practical Machine Learning Course Project - Human Activity Recognition
===
The goal of this project is to predict human activity based on sensor data. We have a training set of 19622 observations and 160 variables , a test set of 20 observations and 160 variables. The target variable is called "classe"" and includes A,B,C,D and E states.


Loading the data
===
```{r,cache=TRUE}
training <- read.csv("pml-training.csv", na.strings=c(".", "NA", "", "?"),header=TRUE, strip.white=TRUE)
testing <- read.csv("testing3.csv", na.strings=c(".", "NA", "", "?"), header=TRUE, sep=",", strip.white=TRUE)
```

Data Munging
===



Part 1- removing ident(X variable)

```{r}
ident.index <- 1

training1 <- training[,-ident.index]
testing1  <- testing [,-ident.index]
names(training)[ident.index]
```



Part 2- Removing columns that have only one unique value

```{r}
factor1.index <- c() 
for (m in (1 : ncol(training1)))   {
  
  
  if (   is.factor(training1[ ,m])  & nlevels(training1[ ,m]) == 1   )
  {   factor1.index <- append(factor1.index,m) 
  }}

training2 <- training1[,-factor1.index]
testing2  <- testing1 [,-factor1.index]
names(training2)[factor1.index] # 6 variables
```



Part 3- Removing Columns that have too many NAs(there are some column that have exactly 19216 NAs)

```{r}
missing.index <- c() 
for (n in (1 : ncol(training2)))   {
  
  
  if (   sum(is.na(training2[ ,n])) == 19216   )
    missing.index <- append(missing.index,n) 
  }
training3 <- training2[,-missing.index]
testing3  <- testing2 [,-missing.index]
head(names(training2)[missing.index],10) #94 variables
```



Part 4 - Finding training and test set variable differences

```{r, eval=FALSE}

setdiff(names(training2),names(testing2)) # classe
setdiff(names(testing2),names(training2)) #problem_id

```



Part 5 - adding problem_id to training dataset and classe to testing dataset to make column names the same

```{r}
# adding columns
training3$problem_id <- NA
testing3 $classe     <- NA

#sorting both training and test datasets based on column names
training3       <- training3[,order(names(training3))]
testing3        <- testing3 [,order(names(testing3))]
```



Partitioning the data
===
```{r}
nobs     <- nrow(training3) # 19622 observations 
sample   <- train <- sample(nrow(training3), 0.7*nobs) # 13735 observations
validate <- sample(setdiff(seq_len(nrow(training3)), train), 0.15*nobs) # 2943 observations
test     <- setdiff(setdiff(seq_len(nrow(training3)), train), validate) # 2944 observations
```



Building the random forest model
===
```{r, cache=TRUE}
input <- names(training3)[! names(training3) %in% c("classes","problem_id")]
target <- "classe"

require(randomForest, quietly=TRUE)

rf <- randomForest(classe ~ .,
        data=training3[sample,c(input, target)],
        ntree=500,mtry=7,importance=TRUE,replace=FALSE)
```



random Forest -Predicting on test set and Publishing Results
===
```{r, eval=FALSE}

newdata.rf <- na.omit(testing3[,c(input)])
pr.rf <- predict(rf, newdata.rf ) 

# Extract the relevant variables from the training3.
sdata.rf <- subset(testing3[,], select=c( "classe"))

# Output the combined data.
write.csv(cbind(sdata.rf, pr.rf), file="testing3_score_idents.csv", row.names=FALSE)

```



Buiding The Second Model for Prediction - Support Vector Machines
===
Using the random forest models 17 out of 20 test cases on the course website passed. For predicting the remaining 3 I used support vector machines.

```{r, eval=FALSE}

require(kernlab, quietly=TRUE)

ksvm <- ksvm(as.factor(classe) ~ .,
      data=training3[training3,c(input, target)],
      kernel="rbfdot",
      prob.model=TRUE)
```



Support Vector Machines -Predicting on test set and Publishing Results
===
```{r,eval=FALSE}

newdata.svm =na.omit(testing3[,c(input)]
# Obtain probability scores for the SVM model on training3.csv.
pr.svm <- predict(ksvm,newdata.svm ))

# Extract the relevant variables from the dataset.
sdata.svm <- subset(testing3[,], select=c("X", "classe"))

# Output the combined data.
write.csv(cbind(sdata.svm, pr.svm), file="C:\Users\Asus\Documents\training3_score_svm.csv", row.names=FALSE)

```



